---
title: "July 7, 2020"
author: "pjaselin"
date: "7/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project Setup

## Library Import
```{r}
library(dplyr)
library(ggplot2)
```

## Data Import
```{r}
coffee_ratings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv')
```

## Helper Functions
```{r}
drop_singular_columns <- function(x) {
  # takes a data frame and returns a data frame without columns that have the same value
  x[vapply(x, function(y) length(unique(y)) > 1, logical(1L))]
}
is.Date <- function(x){
  # takes column or value and checks to see if it is a date --> could be more generalized
  inherits(x, "Date")
}
num_unique <- function(x){
  # gets the number of unique elements in vector x
  length(unique(x))
}
```


```{r}
dim(coffee_ratings)
```



# Study Missingness
```{r}
missingness <- coffee_ratings %>%
  tidyr::gather(key, value) %>%
  mutate(ismissing = is.na(value)) %>%
  group_by(key, ismissing) %>%
  summarise(num_missing = sum(ismissing)) %>%
  filter(ismissing == TRUE) %>%
  mutate(perc_missing = num_missing/dim(coffee_ratings)[1]) %>%
  mutate(perc_present = 1 - perc_missing) %>%
  rename(varname = key)
```

Which columns are complete?
```{r}
setdiff(names(coffee_ratings), missingness$varname)
```


Graph the counts of missing values by column
```{r}
missingness %>%
  ggplot(aes(x = reorder(varname, -perc_missing), y = num_missing)) + 
  geom_col() +
  theme_bw() + 
  theme(axis.text.y = element_text(size = 6),
        plot.title = element_text(hjust = 0.5)) +
  xlab("Column Name") +
  ylab("Missing Value Count") +
  ggtitle("Counts of Missing Values across Columns where they Occur") +
  coord_flip() +
  scale_y_continuous(breaks=scales::pretty_breaks(n=10))
```


Graph the percent missing 
```{r}
perc_missingness <- missingness %>%
  tidyr::gather(key, value, -varname, -ismissing, -num_missing) %>%
  mutate(key = recode(key, "perc_present" = "% Present", "perc_missing" = "% Missing")) %>%
  mutate(key = factor(key, levels = c("% Present", "% Missing"))) 

perc_missingness %>%
  ggplot(aes(x = reorder(varname, -num_missing), y = value, fill = key)) + 
  geom_col() +
  theme_bw() + 
  theme(axis.text.y = element_text(size = 6),
        plot.title = element_text(hjust = 0.5)) +
  xlab("Column Name") +
  ylab("Missing Value Percent [%]") +
  ggtitle("Percentage of Missing Values across Columns where they Occur") +
  coord_flip() +
  scale_y_continuous(breaks=scales::pretty_breaks(n=10)) + 
  guides(fill=guide_legend(title="")) +
  scale_fill_manual(values = c("steelblue", "tomato3"))# +
  #geom_text(aes(label = round(value,4)*100), size = 2)
```




```{r}
reorder_score <- c()
for (score in missingness$num_missing) {
  reorder_score <- c(reorder_score, rep(score, (length(coffee_ratings$total_cup_points)*length(missingness$varname))/length(unique(missingness$varname))))
}

coffee_ratings %>%
  select(missingness$varname) %>%
  mutate(id = row_number()) %>%
  tidyr::gather(-id, key = "key", value = "val") %>%
  mutate(isna = is.na(val)) %>%
  ggplot(aes(reorder(key,-reorder_score), id, fill = isna)) +
  geom_raster() +
  scale_fill_manual(name = "",
                    values = c('steelblue', 'tomato3'),
                    labels = c("Present", "Missing")) +
  theme_bw() +
  theme(axis.text.y = element_text(size = 6),
        plot.title =  element_text(hjust = 0.5)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n=7)) +
  xlab("Variable") +
  ylab("Row Number") + 
  ggtitle("Missing Values in Rows where Missingness Occurs") +
  coord_flip()
```

get the percent missing row by row
```{r}
coffee_ratings %>%
  mutate(id = row_number()) %>%
  tidyr::gather(-id, key = "key", value = "val") %>%
  select(-key) %>%
  group_by(id) %>%
  summarise(perc_na = sum(is.na(val))/length(val)) %>%
  ggplot(aes(x = id, y = perc_na)) +
  geom_line() +
  theme_bw()
```


# Data Cleaning
```{r}
coffee_ratings <- coffee_ratings %>%
  mutate(grading_date = stringr::str_replace(grading_date, "rd,", ",")) %>%
  mutate(grading_date = stringr::str_replace(grading_date, "nd,", ",")) %>%
  mutate(grading_date = stringr::str_replace(grading_date, "th,", ",")) %>%
  mutate(grading_date = stringr::str_replace(grading_date, "st,", ",")) %>%
  mutate(grading_date = stringr::str_replace(grading_date, "\n", "")) %>%
  mutate(grading_date = lubridate::mdy(grading_date)) %>%
  mutate(expiration = stringr::str_replace(expiration, "rd,", ",")) %>%
  mutate(expiration = stringr::str_replace(expiration, "nd,", ",")) %>%
  mutate(expiration = stringr::str_replace(expiration, "th,", ",")) %>%
  mutate(expiration = stringr::str_replace(expiration, "st,", ",")) %>%
  mutate(expiration = stringr::str_replace(expiration, "\n", "")) %>%
  mutate(expiration = lubridate::mdy(expiration))
```



## Value Counts for Categorical Features

```{r}
head(coffee_ratings)
```

how many unique levels are there in each of the discrete columns?
```{r}
coffee_ratings %>%
  select(species, owner, country_of_origin, farm_name, mill, ico_number, company, altitude, producer, in_country_partner, harvest_year, owner_1, variety,
         processing_method, color, category_one_defects, category_two_defects, quakers, certification_body, unit_of_measurement) %>%
  tidyr::gather(key, value) %>%
  group_by(key) %>%
  summarise(num_unique = num_unique(value)) %>%
  arrange(-num_unique) %>%
  mutate(unique_frac = num_unique/nrow(coffee_ratings))
```

let's ignore columns that have high cardinality (above 20%)


```{r, fig.width = 6, fig.height = 4}
coffee_ratings %>%
  select(species, country_of_origin, in_country_partner, variety, processing_method, color, category_one_defects, category_two_defects, quakers, certification_body, unit_of_measurement) %>%
  tidyr::gather(key, value) %>%
  #filter(!is.na(value)) %>%
  ggplot(aes(x = value, width = 0.5)) +
  geom_bar(stat="count", position = "identity") +
  theme_bw() + 
  theme(axis.text.y = element_text(size = 6),
        strip.text.x = element_text(size = 8)) +
  facet_wrap(~key, scales = "free") +
  #scale_y_continuous(breaks=scales::pretty_breaks(n=5)) +
  coord_flip() +
  scale_x_discrete(label=function(x) abbreviate(x, minlength = 7))
```



grading by dates
```{r}
coffee_ratings %>%
  select(grading_date, aroma, flavor, aftertaste, body, balance, uniformity, clean_cup, sweetness, cupper_points, moisture) %>%
  tidyr::gather(key, value, -grading_date) %>%
  ggplot(aes(x = grading_date, y = value)) +
  facet_wrap(~key, scales = "free") +
  geom_point(size=1, shape=16) +
  theme_bw() 
```

grading distributions
```{r}
coffee_ratings %>%
  select(aroma, flavor, aftertaste, body, balance, uniformity, clean_cup, sweetness, cupper_points, moisture) %>%
  tidyr::gather(key, value) %>%
  ggplot(aes(x = key, y = value)) +
  geom_boxplot() +
  theme_bw() 
```


compare reviewers
predict score?
